RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-07-25 overview.tex
metamodel

\todo{Marc}{all}{This section is copy/pasted from RENext, we should revise the metamodel and the text.}

Figure~\ref{fig:metamodel} depicts the metamodel linking the main elements of our argumentation extension to the main GRL elements. The part below the dashed horizontal line depicts GRL elements. A \textsf{GRL Diagram} (bottom) contains zero or more \textsf{IntentionalElements}, either a \textsf{Goal}, a \textsf{Softgoal}, a \textsf{Task}, or a \textsf{Belief}. An \textsf{ElementLink} is either a \textsf{Contribution}, a \textsf{Decomposition}, or a \textsf{Dependency} and contains an \textsf{IntentionalElements} as source and as target.

The part above the horizontal dashed line depicts the concepts we introduced in the previous section. The top-left element \textsf{Argument} (see Section~\ref{sect:arguments}) can attack other arguments (Section~\ref{sect:attacks}) and generalizes both a \textsf{Formula} and an \textsf{Inference}. That is, both these elements can be arguments. A formula has an \textsf{AcceptStatus} (Section~\ref{sect:extensions}). An \textsf{Inference} is from a set of \textsf{Arguments} as premise and a \textsf{Formula} as conclusion. The \textsf{InferenceType} is either strict or defeasible (Section~\ref{sect:aspic}). A \textsf{Formula} is either a \textsf{Modality} (modal formula), a \textsf{Proposition}, a \textsf{BinaryOperation}, or a \textsf{Negation} (negated proposition). The \textsf{Modality} can be either B (belief), G (goal), E (evidence), or A (action). A \textsf{BinaryOperation} is either a \textsf{Disjunction}, a \textsf{Conjunction}, or a \emph{contributes\_to} \textsf{Connective}~(\ref{sect:aspic}).

Finally, the red arrows depict how the two metamodels are integrated. The left arrow connects a GRL \textsf{IntentionalElement} with an argumentation \textsf{Modality}, where the mapping is denoted with red text. Thus, an intention element traces to an argument, which is always a modal formula. The right arrow connects the GRL \textsf{ElementLink} with the argumentation \textsf{Connective} (i.e. a \emph{contributes\_to} connective). Thus, an arrow between elements of a GRL model corresponds to \emph{contributes\_to} connectives in the argumentation framework.

\begin{figure*}[h!]
\includegraphics[scale=0.61]{metamodel/metamodel}
\caption{The RationalGRL metamodel}
\label{fig:metamodel}
\end{figure*}

>>>>

Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements. We describe the metamodel bottom-up, starting with the GRL package.

The GRL package of our metamodel consists of the core GRL concepts, which constitute a part of
the URN metamodel from Recommendation Z.151~\cite{URN}. These concepts represent the abstract grammar of the language, independently of the notation. This metamodel also formalizes the GRL concepts and constructs introduced earlier\footnote{Note that for readability, some GRL concepts ave been omitted from the figure. For instance, a GRL contribution can have a qualitative strength, ranging from ``Make'' to ``Break''. Since these concepts are not relevant to our framework, they have been omitted, but none of the results of this article depend on this.}.

The GRL specification consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}. A \textsf{GRLLinkableElement} can again be specialized into an \textsf{Actor} or an \textsf{IntentionalElement} (which is either a \textsf{Softgoal}, \textsf{Goal}, \textsf{Task}, \textsf{Resource}, or a \textsf{Belief}). Intentional elements can be part of an actor, and \textsf{GRLLinkableElements} are connected through \textsf{ElementLinks} of different types (i.e., \textsf{Contribution, Decomposition}, or \textsf{Dependency}). Note that actors can be connected through links as well, which is done wth \textsf{Dependency} links.

The Argumentation package depicts the concepts we introduced in the previous sections. An \textsf{ArgumentScheme} represents an (uninstantiated) scheme containing variables that can be replaced with intentional elements. \textsf{CriticalQuestions} are possible ways to attack or elaborate an argument scheme. As such, each critical question applies to exactly one scheme, but each scheme can be elaborated or attack through multiple critical questions. When an argument scheme is instantiated, we obtain an \textsf{Argument}. Therefore, each argument is associated with exactly one scheme, but a scheme can be instantiated in multiple ways. When a critical question is answered, we obtain an \textsf{AttackLink}. Each \textsf{AttackLink} is associated with at most one critical question, but a critical question can be used to attack multiple arguments.  Note that an \textsf{AttackLink} can also be associated with no critical questions. This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions. A \textsf{RationalGRLDiagram} is composed out of arguments and attack relations.

There is only one link between the \textsf{GRL} package and the \textsf{Argumentation} package, but it is a very important one. The link specifies that each \textsf{GRLModelElement} is in fact an argument. This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted, rejected, or undecided. This furthermore means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme. Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements, but can come from an external source, for instance a piece of evidence or an expert opinion. We will see various examples of such arguments in the next sections.

\begin{figure*}[h!]
\includegraphics[width=\textwidth]{metamodel/metamodel}
\caption{The RationalGRL metamodel}
\label{fig:metamodel}
\end{figure*}
